Semantic prediction in {R}ussian

### 2020-04-10 paper.tex
Updates from Overleaf

\documentclass[a4paper]{article}
\usepackage{lrec}
\usepackage{amsmath}
\usepackage{polyglossia}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue,
}

@@

% TODO (@SEMEN SOROKIN) объединить с интродакшн, мб как-то связать или убрать вообще, т.к. абзац незаконченный.
\section{Background}

% TODO: more background required; let's include our reviews from the
% bibliography N. Lopukhina provided us?

As Kuperberg and Jaeger claim in their “What do we mean by prediction
in language comprehension?” \cite{kuperberg-jaeger}, the reaction time is in direct ratio with the predictability of the word: the more predictable the word is the faster is the reaction time. Moreover, predictability of a word or a context
% TODO [f]
  defines
% TODO [g]
 fixation time in eye-movement studies as a result of language comprehension process. This implies that language comprehension must
% TODO [h]
 be predictive. The authors also state that as the previous context expands, the predictability of the next word increases leading to - in cloze tests - higher accuracy of predicting the next work, and - in eye-movement experiments - to shorter fixation duration
% TODO [i]
.

\section{Related works}
% TODO (@Anastasia Nikiorova) Add article reviews for related works

There are several scientific papers that propose different algorithms for cloze answer generation. In \cite{zhou-cloze} the authors state the importance of next word prediction in language modelling and its potential contribution to OCR and hand writing recognition. The authors enhance existing models with ELMo and BERT language models and train on the CLOTH dataset of cloze tests. Expectedly, BERT models show the highest performance (0.86 and 0.83 accuracy scores on test dataset for BERT Large and BERT Base respectively), as this model was initially trained to recover masked tokens in text. At the same time, ELMo model's poor performance could be due to the lack of parameter tuning and the fact that ELMo was trained for next sequence word prediction.
An LSTM-based model for cloze-style machine comprehension is proposed in \cite{wang-cloze}. The model consists of document hierarchical structure and dynamic attention mechanism for building the representations between the document and the question.

@@

% TODO begin by defining a cloze task.
Cloze task described in Taylor (1953) is an experiment in which one or more words are removed from a sentence and the participants are asked to fill in the missing content. 
We used the dataset with cloze task answers from the study
\cite{laurinavichyute}. This dataset is based on 144 sentences
randomly selected from the Russian National Corpus (RNC,
 ruscorpora.ru) - an online corpus of Russian texts with extensive
search options. These sentences were slightly edited: the authors
replaced rare infrequent words with more frequent ones and shortened
the sentences when they exceeded the preset maximum length of 13
words. The stimuli sentences were subjected to the cloze task
experiment. Cloze task \cite{taylor:1953} is a test, commonly carried out
for assessing native speakers of a language, which is aimed to
understand respondents’ comprehension of a language and their ability
to predict missing portions of written texts \cite{laurinavichyute}. This experiment presumes that native speakers are able to understand context and vocabulary in order to identify the correct semantic field or part of speech of a missing word.
% TODO [j]
 In \cite{laurinavichyute}, the respondents were asked to successively predict next words for each context. An example of stimuli that were shown to the participants is presented in the table below with corresponding correct next words and calculated predictability scores.

\begin{table*}
\centering
\begin{tabular}{llr}

@@

% TODO [k]

@@

A one-layer long short-term memory (LSTM) recurrent neural network model was used
\cite{jozefowicz:2016}
%[Jozefowicz, Vinyals, Schuster, Shazeer, Wu, 2016]
to create a list of predictions for each word in the same 144 stimuli sentences. The dimensions of the model are 2048 for hidden layer and 512 for the input and output layers.

@@

Bidirectional Encoder Representations from Transformers
(BERT) \cite{bert} is trained on a masked language modeling objective. Unlike a traditional language modeling objective of predicting the next word in a sequence given the history, masked language modeling predicts a word given its left and right context. Because the model expects context from both directions, it is not obvious how BERT can be used as a traditional language model (i.e. to evaluate the probability of a text sequence) or how to sample from it. We test several ideas: give the model all of content, except masked word, and using the technique[] to rework BERT as classical language model.
% TODO [l]

For experiments we will using BERT trained on Russian Wiki corpus
\cite{ruwiki}. To show differences  between models, we will fine-tune BERT with our corpuses.
% TODO [m]

% TODO [n]

@@

% TODO: citation for russian news corpus
Russian news corpus \cite{runews} includes newspaper articles published in the 2000s: newspaper different types of articles.
% TODO [p]

% TODO: citation for natcorp
National corpus of Russian language \cite{natcorp} include written texts from the middle of the 18th to the middle of the 20th centuries
% TODO [q]
.

@@

\begin{table}
\caption{Corpus statistics. RNC is the Russian News Corpus, and NCRL
is the National Corpus of the Russian Language.}
\centering

@@

\begin{itemize}
	\item Hidden Markov Models
	\begin{itemize}
		\item Bigram HMM on RNC
		\item Bigram HMM on NCRL
		\item Trigram HMM on RNC
		\item Trigram HMM on NCRL
	\end{itemize}
	\item BERT-based models
% TODO [r]
	\begin{itemize}
		\item no fine-tuning
		\item BERT fine-tuned on RNC
		\item BERT fine-tuned on NCRL
		\item BERT fine-tuned on both
	\end{itemize}
	% TODO decide whether the word "custom" is appropriate here;
	% I think BERT also uses LSTMs, so probably this is just "an
	% LSTM model of our own design"?
	\item custom LSTM model
% TODO [s]
	\item hybrid model
	\begin{itemize}
		\item Bigram HMM on NCRL + BERT fine-tuned on NCRL
	\end{itemize}
\end{itemize}

LSTM was trained on the part of RNC
% TODO [t]
, excluding the sentences chosen for stimuli, namely on modern written texts from the 1950s to the present day, real-life Russian oral speech recordings from the same period. In addition
% TODO [u]
, it was trained on RNC and NCRL. Overall, the training corpus consisted of 577 million tokens. The model was tested on 1000 sentences from the OpenCorpora 
\cite{bocharov:2011}
% [http://opencorpora.org/; Bocharov
% TODO [v]
% et al. 2011]
with overall 1,9 million of tokens from newspaper articles, Russian Wikipedia, texts from blogs, fiction, non-fiction, and legal documents.

Renormalization of probabilities

@@

% TODO [w]

@@

The metric represents the amount of contexts for each prediction of the correct word is non-zero.    
% TODO [x]

@@

% TODO [y]

@@

\[ {D_{n,m}=\sup _{x}|F_{1,n}(x)-F_{2,m}(x)|,} \]
% TODO [z]

@@

This metric was used to compare both lexical and word class
% TODO [aa]
 probability distributions of cloze task, LSTM, HMM models pairwise. The results of the metric are listed in the Results section of this article.

\head{Kullback–Leibler divergence} % Sergey

Cross-entropy loss, or log loss, measures the performance of a classification
% TODO [ab]
 model whose output is a probability value between 0 and 1. Cross-entropy loss increases as the predicted probability diverges from the actual label. The cross-entropy shows the difference between probability distributions p and q. Kullback and Leibler defined a similar measure now known as KL divergence. This measure quantifies how similar a probability distribution p is to a candidate distribution q. We used KL-divergence to compare several
% TODO [ac]
 language model.

\subsection{PoS} % [ad][ae]  probabilities Daria

Word
% TODO [af]
 class probabilities were computed as follows: each word in the model’s
% TODO [ag]
 vocabulary (\(N = 500000\) most frequent words in the training corpus) and each word in the stimuli sentences was tagged for word class and morphological features using PyMorphy2 analyser 
\cite{korobov:2015}
%[Korobov
% TODO [ah]
%, 2015]
and the predictions of the model were compared to the
annotation of the target words. A word class match was coded if the
predicted and target word belonged to the same word class. The
probability of a word class was computed by summing probabilities of
all words in the model’s vocabulary which had the morphosyntactic
feature in question. For example, to estimate the word class
probability in a sentence “A mobile \underline{\hspace{1em}}”, where "phone" is the target word, we would sum up probabilities of all nouns in the model’s vocabulary. For morphologically ambiguous words (e.g., рот ‘mouth’ in nominative or accusative singular), all possible variants were considered in the probability estimation.

\subsection{Probabilities for OBJECT-VERB-FUNCTIONAL-MODIFIER} % Daria

@@

Firstly, we converted all of the modified Pymorphy tags into 4 general sets: 'ADJ', 'ADVB', 'NUMR' were generalized to 'MOD'; 'INFN' and 'PRED' - to 'VERB'; 'NPRO' to 'NOUN'; 'PREP', 'PRCL', 'CONJ' and 'INTJ' to 'FUNC' for each context. Then, we have counted probabilities of these tags in the same manner as the usual parts of speech
% TODO [ai]
.

@@

\subsection{Semantic comparison} % Daria

After lots of trial and error
% TODO [aj]
, it was decided to use semantic vectors for the comparison of cloze task with other models in the semantic aspect
% TODO [ak]
. All words were mapped into a vector space of the model pre-trained on Wikipedia texts
% TODO [al]
 \cite{arefyev}. The comparison itself needed to be dynamic, because for each context a different amount of words should have been chosen.

To compare semantic vectors for each context, firstly, we cleared all of the words so there would be no digits, meaningless letters and punctuation. Then, we have built a function
% TODO [am]
, which:

1)        Extracts first 10 words (we have decided that that is the maximum amount of words for each context that would have meaningful probabilities as all of the words after the first ten for each model have their probabilities tends to zero) for each of 1219 contexts
% TODO [an]
;

@@

3)        Counts the difference between the probabilities of the first and last word
% TODO [ao]
 and then difference between mean probabilities of the previous word and the next one
% TODO [ap]
;

@@

Then, MiniBatchKmeans (a modified version of the K-means algorithm,
that uses mini-batches to reduce computation time, while at the same
time trying to optimize the same goal function \cite{bejar}) algorithm in sklearn was used to find “cluster centers”, or mean semantic vector for one model for each context
% TODO [aq]
.
%And at last, we found cosine similarity (also with sklearn library) for each pair of semantic vectors for each pair of models.

%\subsection{Cosine similarity}
%
%Cosine similarity was used to measure closeness of semantic vectors of predicted words between different models.
%
%Cosine similarity is a measure of similarity between two non-zero vectors that measures the cosine of the angle between them. Given two vectors of attributes, A and B, the cosine similarity, cos(θ), is calculated as
%
%
%
%\[ {{\text{similarity}}=\cos(\theta )=\frac{{A} \cdot {B}}{\|{A}
%\|\|{B} \|}={\frac {\sum \limits _{i=1}^{n}{A_{i}B_{i}}}{{\sqrt {\sum
%\limits _{i=1}^{n}{A_{i}^{2}}}}{\sqrt {\sum \limits
%_{i=1}^{n}{B_{i}^{2}}}}}},}  \]
%
%where \({A_{i}}\) and \({B_{i}}\) are components of vector {A} and {B} respectively.
%
%The cosine similarity is widely used for calculating distance between two words. In our study, cosine similarity was used to find semantic probability of predicting the word which is semantically close to the target word.

@@

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{figures/pdf/num-fill-context-dist.pdf}
\end{figure}
% TODO [ar]

@@

In Table~\ref{tbl:min-var}, contexts with minimum variance in filling
are listed. It is worth emphasizing that in all the above cases there
was no variance in respondents’ answers, i.e. all respondents gave the
same one answer for these contexts. What is more, this predicted word
was the original word from the corpus.

We classified these contexts based on their constraining ability
% TODO [as]
 and found out the following groups of com:

@@

\begin{figure}
\caption{Context length vs. accuracy.}
\centering

\includegraphics[width=0.4\textwidth]{figures/pdf/context-accuracy.pdf}

\end{figure}

@@

In Figure~\ref{fig:context-accuracy-regression} reflects a high value of predictability for contexts of length both less and more than five lexical units. However, the lack of correlation is worth emphasizing. We received Pearson correlation coefficient score of 0.323 and Spearman correlation coefficient of 0.363.

According to the results of our experiment, the closest probability distribution of the correct word of all the models was achieved with BERT trained on the literary
% TODO [at]
 corpus.

In this case
% TODO [au]
, there is practically no correlation
% TODO [av]
: Pearson correlation score is 0.093521, Spearman correlation - 0.079094.

@@

% TODO [aw]

Each model was evaluated by two measures: mean accuracy of model predictions and absolute number of correct word predictions. For computing mean accuracy, the mean of correct answer probabilities was taken. In case of absolute number of correct word predictions, a model achieved +1 score if there was at least one correct answer among all predictions.
% TODO [ax]

@@

Figure~\ref{fig:mean-accuracy-hist-lex}
% TODO [ay]
% TODO [az]
 below shows a bar chart of the mean accuracy scores of each model on
the lexical level. As the goal of our study was to build an algorithm,
which would be the closest approximation of the cloze task results
(18\% accuracy), we can see that BERT (not a language model one) model
scored better than the others. Interestingly, all HMM model
architectures showed low results on the lexical level.

@@

It is noticeable that BERT mean accuracy results are higher than cloze task score. This can be explained by the fact that the model was trained on a large amount of written texts and thus had a higher chance to guess the correct word. Following this assumption, an assumption can be made that respondents’ active vocabulary size is lower than model’s vocabulary. Also, we can make a hypothesis that the process of word retrieval by humans and by the model is performed differently, as respondents do not always respond with the most probable answer
% TODO [ba]
.

@@

At the next step, we performed Kolmogorov-Smirnov testing to find the similarity in the probability distributions of the model predictions. This table also reflects the sum value of pairwise comparison of the contexts using Kolmogorov-Smirnov testing. For convenience all the values were normalized. Observations show that the closest probability distribution is seen in LSTM model. However, as we see, there is a variation in metric values from 0.9 to 6.2 for different models. Thus, we can assert that the difference between all models and a cloze task is rather large and in some way unacceptable.
% TODO [bb]

%\head{Cosine Similarity}
%To measure model predictions on semantic level, the cosine similarity between each context’s predicted words centroid vector was found. The number of words was selected dynamically for each context by maximizing vector significance with the minimum words.
%

% TODO [bc]

\head{Kullback–Leibler divergence} % Sergey

Due to the fact that all of our models are word-level, and in order to lower the casing variability we've combined all our vocabularies into one
% TODO [bd]
. For this compound vocabulary we calculated the KL divergence of our models.

The table below shows the scores for three different models with
bigram HMM trained on NCRL showing the best results. Unfortunately,
this heat map does not show us changes in LM \{language model?\} distances context-lengthwise. Table of top-3 of lengthwise LM distances.

@@

\caption{KL scores for three different models, all trained on the
National Corpus of the Russian Language. ``Context provided'' is in number
of tokens.}

@@

\subsection{Model comparison on the PoS level} % Semen, Anastasia

@@

The performance of all models, except for LSTM
% TODO [be]

% TODO [bf]

% TODO [bg]
, at the parts of speech level has significantly and proportionally increased. This is due to a decrease in the set of classes for which classification occurs. At this stage, the first 30 words of each model were tagged for parts of speech. Overall, there were 16 word classes. It is notable that BERT linguistic models have the highest scores.

@@

\subsection{Model comparison on the OMVF level} %  Semen, Daria

@@

Reducing the number of classes by 4 times does not lead to an improvement in the average accuracy
% TODO [bh]
. Although we can see that the models perform similar as on the original parts of speech model.

@@

\head{Kolmogorov-Smirnov}%KS test (based on absolute value statistics)

@@

\section{Conclusion}

\section{Bibliography}% [bi] 
\bibliographystyle{lrec}
\bibliography{paper}

@@

% [a]no abstract
% 
% [b]it is very limited in the text
% 
% [c]it is not present in the text...
% 
% [d]task?
% 
% [e]not clear, how probability score is calculated
% 
% [f]??? can we measure predictability of a context?
% 
% [g]too strong
% 
% explains variance
% 
% [h]other papers show that it is not an oblagation
% 
% [i]incorrect implication
% 
% [j]does it?
% 
% [k]what is the rationale behind all these models?
% 
% Why did you choose them?
% 
% [l]not clear: how did you plan to retransform BERT?
% 
% [m]how this sentence is related to the previous one?
% 
% [n]Непонятно: в итоге BERT у вас видит контекст справа или нет?
% 
% [o]why did you select these particular corpora?
% 
% [p]link
% 
% [q]and what about later texts?
% 
% [r]what corpus it was trained on?
% 
% [s]trained on what?
% 
% [t]did you train your own lstm?
% 
% [u]why?
% 
% [v]add to the references
% 
% [w]how was it computed
% 
% [x]how was it computed
% 
% [y]how was it computed
% 
% [z]latex format for the final article
% 
% [aa]what is it? it was not introduced above
% 
% [ab]which of the models was a classification model? it was not explained above
% 
% [ac]unclear
% 
% [ad]what it the rationale to compute Pos probabilities?
% 
% [ae]PoS or word class? choose one nomination
% 
% [af]it is my text
% 
% [ag]did you do this for the cloze task responses?
% 
% [ah]references
% 
% [ai]did you do this for models' predictions or for cloze responses as well?
% 
% [aj]what did you try? can you list it?
% 
% [ak]what does it mean? what was the goal?
% 
% [al]why did you choose this model???
% 
% [am]какова идея за этой функцией? вы можете сначала объяснить, что вы в принципе хотели сделать?
% 
% [an]how did you get this number?
% 
% [ao]where? in the context?
% 
% [ap]in context?
% 
% [aq]for each context or for each word in context?
% 
% why did you measure semantic similarity of contexts?
% 
% [ar]number of guesses for each target word?
% 
% [as]More info on constraining contexts with reference to the original article
% 
% [at]which one?
% 
% [au]not clear
% 
% [av]between what and what?
% 
% [aw]do this figures illustrate BERT??
% 
% [ax]not clear
% 
% [ay]I am surprised with the lstm results
% 
% In our study we got 19% accuracy for lstm trained on the national corpus of Russian
% 
% [az]https://colab.research.google.com/drive/1vZ4a2T-cCpE76VGvw4E-IyODy7XapWkc Добрый день, здесь выложил часть кода, отвечающую за расчет метрики для ЛСТМ,  не могли бы сверить со своими результатами?
% 
% [ba]это довольно странное заключение: что мы считаем наиболее вероятным ответом? Что у нас является золотым стандартом вероятности ответа? Почему ответ BERT более вероятный, чем ответ людей?
% 
% [bb]why? what are the acceptability criteria?
% 
% [bc]what does the heatmap reflect?
% 
% [bd]why did you decide to do this?
% 
% [be]this is a really strange result
% 
% Can you double-check it?
% 
% [bf]and it contradicts your calculations below (3)
% 
% [bg]Непонятно, как PoS accuracy вообще может быть ниже word accuracy: если модель правильно предсказала слово, то она и его часть речи правильно предсказала.
% 
% [bh]this is also very strange
% 
% Are you absolutely sure about these results? How can you explain them?
% 
% [bi]should be aligned with the text

>>>>

\documentclass[a4paper]{article}
\usepackage{lrec}
\usepackage{amsmath}
\usepackage{polyglossia}
\usepackage{graphicx}
\usepackage{hyperref}

@@

\begin{table*}
\caption{Example of stimuli sentence from RNC in cloze task with probabilities of the next word. Sentence: "А промывать манную крупу перед тем, как варить ее, не пробовали?" (English translation: "Have you tried to rinse semolina before boiling it?")
\newline}
\centering
\begin{tabular}{llr}

\section{Related works}

As Kuperberg and Jaeger claim in their “What do we mean by prediction
in language comprehension?” \cite{kuperberg-jaeger}, the reaction time is in direct ratio with the predictability of the word: the more predictable the word is the faster is the reaction time. Moreover, predictability of a word or a context defines fixation time in eye-movement studies as a result of language comprehension process. This implies that language comprehension must be predictive. The authors also state that as the previous context expands, the predictability of the next word increases leading to - in cloze tests - higher accuracy of predicting the next work, and - in eye-movement experiments - to shorter fixation duration.

There are several scientific papers that propose different algorithms for cloze answer generation. In \cite{zhou-cloze} the authors state the importance of next word prediction in language modelling and its potential contribution to OCR and hand writing recognition. The authors enhance existing models with ELMo and BERT language models and train on the CLOTH dataset of cloze tests. Expectedly, BERT models show the highest performance (0.86 and 0.83 accuracy scores on test dataset for BERT Large and BERT Base respectively), as this model was initially trained to recover masked tokens in text. At the same time, ELMo model's poor performance could be due to the lack of parameter tuning and the fact that ELMo was trained for next sequence word prediction.

An LSTM-based model for cloze-style machine comprehension is proposed in \cite{wang-cloze}. The model consists of document hierarchical structure and dynamic attention mechanism for building the representations between the document and the question. Despite the two-layer LSTM model with attention outperforms one-layer model, the final best accuracy score is still only 0.76 which could be improved by future modifications to the model.

@@

Cloze task described in Taylor (1953) is an experiment in which one or more words are removed from a sentence and the participants are asked to fill in the missing content. 
We used the dataset with cloze task answers from the study
\cite{laurinavichyute}. This dataset is based on 144 sentences
randomly selected from the Russian National Corpus (RNC,
 ruscorpora.ru) - an online corpus of Russian texts with extensive
search options. These sentences were slightly edited: the authors
replaced rare infrequent words with more frequent ones and shortened
the sentences when they exceeded the preset maximum length of 13
words. The stimuli sentences were subjected to the cloze task
experiment. Cloze task \cite{taylor:1953} is a test, commonly carried out
for assessing native speakers of a language, which is aimed to
understand respondents’ comprehension of a language and their ability
to predict missing portions of written texts \cite{laurinavichyute}. This experiment presumes that native speakers are able to understand context and vocabulary in order to identify the correct semantic field or part of speech of a missing word.

In \cite{laurinavichyute}, the respondents were asked to successively predict next words for each context. An example of stimuli that were shown to the participants is presented in Table 1 with corresponding correct next words and calculated predictability scores.

@@

@@

A one-layer long short-term memory (LSTM) recurrent neural network model was used \cite{jozefowicz:2016} to create a list of predictions for each word in the same 144 stimuli sentences. The dimensions of the model are 2048 for hidden layer and 512 for the input and output layers.

@@

Bidirectional Encoder Representations from Transformers
(BERT) \cite{bert} is trained on a masked language modeling objective. Unlike a traditional language modeling objective of predicting the next word in a sequence given the history, masked language modeling predicts a word given its left and right context. Because the model expects context from both directions, it is not obvious how BERT can be used as a traditional language model (i.e. to evaluate the probability of a text sequence) or how to sample from it. We test several ideas: give the model all of content, except masked word, and using the technique[] to rework BERT as classical language model.

For experiments we will using BERT trained on Russian Wiki corpus
\cite{ruwiki}. To show differences  between models, we will fine-tune BERT with our corpuses.

@@

Russian news corpus \cite{runews} includes newspaper articles published in the 2000s: newspaper different types of articles.
National corpus of Russian language \cite{natcorp} include written texts from the middle of the 18th to the middle of the 20th centuries.

@@

\begin{itemize}
	\item Hidden Markov Models
	\begin{itemize}
		\item Bigram HMM on RNC
		\item Bigram HMM on NCRL
		\item Trigram HMM on RNC
		\item Trigram HMM on NCRL
	\end{itemize}
	\item BERT-based models

	\begin{itemize}
		\item no fine-tuning
		\item BERT fine-tuned on RNC
		\item BERT fine-tuned on NCRL
		\item BERT fine-tuned on both
	\end{itemize}

	\item custom LSTM model
	\item hybrid model
	\begin{itemize}
		\item Bigram HMM on NCRL + BERT fine-tuned on NCRL
	\end{itemize}
\end{itemize}

LSTM was trained on the part of RNC, excluding the sentences chosen for stimuli, namely on modern written texts from the 1950s to the present day, real-life Russian oral speech recordings from the same period. In addition, it was trained on RNC and NCRL. Overall, the training corpus consisted of 577 million tokens. The model was tested on 1000 sentences from the OpenCorpora 
\cite{bocharov:2011} with overall 1,9 million of tokens from newspaper articles, Russian Wikipedia, texts from blogs, fiction, non-fiction, and legal documents.

\begin{table}
\caption{Corpus statistics. RNC is the Russian News Corpus, and NCRL is the National Corpus of the Russian Language.
\newline}
\centering

@@

\head{Renormalization of probabilities}

@@

@@

The metric represents the amount of contexts for each prediction of the correct word is non-zero.

@@

@@

\[ {D_{n,m}=\sup _{x}|F_{1,n}(x)-F_{2,m}(x)|,} \]

@@

This metric was used to compare both lexical and word class probability distributions of cloze task, LSTM, HMM models pairwise. The results of the metric are listed in the Results section of this article.

\head{Kullback–Leibler divergence}

Cross-entropy loss, or log loss, measures the performance of a classification model whose output is a probability value between 0 and 1. Cross-entropy loss increases as the predicted probability diverges from the actual label. The cross-entropy shows the difference between probability distributions p and q. Kullback and Leibler defined a similar measure now known as KL divergence. This measure quantifies how similar a probability distribution p is to a candidate distribution q. We used KL-divergence to compare several language model.

\head{Cosine similarity}

Cosine similarity was used to measure closeness of semantic vectors of predicted words between different models. It is widely used for calculating distance between two words. In our study, cosine similarity was used to find semantic probability of predicting the word which is semantically close to the target word. 
%
%Cosine similarity is a measure of similarity between two non-zero vectors that measures the cosine of the angle between them. Given two vectors of attributes, A and B, the cosine similarity, cos(θ), is calculated as
%
%
%
%\[ {{\text{similarity}}=\cos(\theta )=\frac{{A} \cdot {B}}{\|{A}
%\|\|{B} \|}={\frac {\sum \limits _{i=1}^{n}{A_{i}B_{i}}}{{\sqrt {\sum
%\limits _{i=1}^{n}{A_{i}^{2}}}}{\sqrt {\sum \limits
%_{i=1}^{n}{B_{i}^{2}}}}}},}  \]
%
%where \({A_{i}}\) and \({B_{i}}\) are components of vector {A} and {B} respectively.

\subsection{PoS}

Word class probabilities were computed as follows: each word in the model’s vocabulary (\(N = 500000\) most frequent words in the training corpus) and each word in the stimuli sentences was tagged for word class and morphological features using PyMorphy2 analyser 
\cite{korobov:2015} and the predictions of the model were compared to the annotation of the target words. A word class match was coded if the predicted and target word belonged to the same word class. The probability of a word class was computed by summing probabilities of all words in the model’s vocabulary which had the morphosyntactic feature in question. For example, to estimate the word class probability in a sentence “A mobile \underline{\hspace{1em}}”, where "phone" is the target word, we would sum up probabilities of all nouns in the model’s vocabulary. For morphologically ambiguous words (e.g., рот ‘mouth’ in nominative or accusative singular), all possible variants were considered in the probability estimation.

\subsection{Probabilities for OBJECT-VERB-FUNCTIONAL-MODIFIER}

@@

Firstly, we converted all of the modified Pymorphy tags into 4 general sets: 'ADJ', 'ADVB', 'NUMR' were generalized to 'MOD'; 'INFN' and 'PRED' - to 'VERB'; 'NPRO' to 'NOUN'; 'PREP', 'PRCL', 'CONJ' and 'INTJ' to 'FUNC' for each context. Then, we have counted probabilities of these tags in the same manner as the usual parts of speech.

@@

\subsection{Semantic comparison}

After lots of trial and error, it was decided to use semantic vectors for the comparison of cloze task with other models in the semantic aspect. All words were mapped into a vector space of the model pre-trained on Wikipedia texts \cite{arefyev}. The comparison itself needed to be dynamic, because for each context a different amount of words should have been chosen.

To compare semantic vectors for each context, firstly, we cleared all of the words so there would be no digits, meaningless letters and punctuation. Then, we have built a function, which:

1)        Extracts first 10 words (we have decided that that is the maximum amount of words for each context that would have meaningful probabilities as all of the words after the first ten for each model have their probabilities tends to zero) for each of 1219 contexts;

@@

3)        Counts the difference between the probabilities of the first and last word and then difference between mean probabilities of the previous word and the next one;

@@

Then, MiniBatchKmeans (a modified version of the K-means algorithm,
that uses mini-batches to reduce computation time, while at the same
time trying to optimize the same goal function \cite{bejar}) algorithm in sklearn was used to find “cluster centers”, or mean semantic vector for one model for each context. And at last, we found cosine similarity (also with sklearn library) for each pair of semantic vectors for each pair of models.

@@

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{figures/pdf/num-fill-context-dist.pdf}
\end{figure}

@@

In Table 3, contexts with minimum variance in filling
are listed. It is worth emphasizing that in all the above cases there was no variance in respondents’ answers, i.e. all respondents gave the same one answer for these contexts. What is more, this predicted word was the original word from the corpus.

We classified these contexts based on their constraining ability and found out the following groups of com:

@@

@@

Regression line reflects a high value of predictability for contexts of length both less and more than five lexical units. However, the lack of correlation is worth emphasizing. We received Pearson correlation coefficient score of 0.323 and Spearman correlation coefficient of 0.363.

According to the results of our experiment, the closest probability distribution of the correct word of all the models was achieved with BERT trained on the literary corpus.

In this case, there is practically no correlation: Pearson correlation score is 0.093521, Spearman correlation - 0.079094.

@@

Each model was evaluated by two measures: mean accuracy of model predictions and absolute number of correct word predictions. For computing mean accuracy, the mean of correct answer probabilities was taken. In case of absolute number of correct word predictions, a model achieved +1 score if there was at least one correct answer among all predictions.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{figures/pdf/context-accuracy-lexical.pdf}
\caption{Context length vs. lexical accuracy.}
\end{figure}

@@

Figure~\ref{fig:mean-accuracy-hist-lex} below shows a bar chart of the mean accuracy scores of each model on
the lexical level. As the goal of our study was to build an algorithm,
which would be the closest approximation of the cloze task results
(18\% accuracy), we can see that BERT (not a language model one) model
scored better than the others. Interestingly, all HMM model
architectures showed low results on the lexical level.

@@

It is noticeable that BERT mean accuracy results are higher than cloze task score. This can be explained by the fact that the model was trained on a large amount of written texts and thus had a higher chance to guess the correct word. Following this assumption, an assumption can be made that respondents’ active vocabulary size is lower than model’s vocabulary. Also, we can make a hypothesis that the process of word retrieval by humans and by the model is performed differently, as respondents do not always respond with the most probable answer.

@@

At the next step, we performed Kolmogorov-Smirnov testing to find the similarity in the probability distributions of the model predictions. Figure 5 also reflects the sum value of pairwise comparison of the contexts using Kolmogorov-Smirnov testing. For convenience all the values were normalized. Observations show that the closest probability distribution is seen in LSTM model. However, as we see, there is a variation in metric values from 0.9 to 6.2 for different models. Thus, we can assert that the difference between all models and a cloze task is rather large and in some way unacceptable.

\begin{figure}
\caption{Kolmogorov-Smirnov tests, lexical level.}
\label{fig:ks-lexical}
\centering
\includegraphics[width=0.5\textwidth]{figures/pdf/ks-lexical.pdf}
\end{figure}

\head{Cosine Similarity}
To measure model predictions on semantic level, the cosine similarity between each context’s predicted words centroid vector was found. The number of words was selected dynamically for each context by maximizing vector significance with the minimum words.

\head{Kullback–Leibler divergence}

Due to the fact that all of our models are word-level, and in order to lower the casing variability we've combined all our vocabularies into one. For this compound vocabulary we calculated the KL divergence of our models.

Table 3 shows the scores for three different models with
bigram HMM trained on NCRL showing the best results. Unfortunately,
this heat map does not show us changes in LM \{language model?\} distances context-lengthwise. Table of top-3 of lengthwise LM distances.

@@

\caption{KL scores for three different models, all trained on the
National Corpus of the Russian Language. ``Context provided'' is in number
of tokens.
\newline}

@@

\subsection{Model comparison on the PoS level}

@@

The performance of all models, except for LSTM, at the parts of speech level has significantly and proportionally increased. This is due to a decrease in the set of classes for which classification occurs. At this stage, the first 30 words of each model were tagged for parts of speech. Overall, there were 16 word classes. It is notable that BERT linguistic models have the highest scores.

@@

\subsection{Model comparison on the OMVF level}

@@

Reducing the number of classes by 4 times does not lead to an improvement in the average accuracy. Although we can see that the models perform similar as on the original parts of speech model.

@@

\head{Kolmogorov-Smirnov Test}

@@

\section{Conclusion}
One of the important results of this study is the development of a certain set of methods (tools) for comparing the generative properties of language models. The starting point is an unrestricted set of contexts and the probabilistic distribution of words for each of them. This data can be obtained from all kind of language models. In addition re-normalized value of frequency from corpus can also be used as a language model for these purposes in further research. From our point of view the most relevant metrics (from presented above) are Mean (Absolute) accuracy is prediction and Cosine similarity measure computed for each pairs of models.  We tested this methodology on the following models: cloze-test-based model, hidden Markov model (with the different n-grams), LSTM, BERT and see that the last one can predict the next word more accurate and LSTM - Cloze task pair shows that semantic directions of k-first words for given context are more similar. However based on all of these metrics scores we can conclude that cloze-test-based model cannot be replaced by any of presented artificial language models for the purposes of eye-tracking experiments and also it is worth noticeable that predictability scores computed on cloze test reflect real-world situation, but last question is beyond the scope of our study.
\section{Bibliography}
\bibliographystyle{lrec}
\bibliography{paper}

@@

% [at]which one?
% 
% [au]not clear
% 
% [av]between what and what?
% 
% [aw]do this figures illustrate BERT??
% 
% [ax]not clear
% 
% [ay]I am surprised with the lstm results
% 
% In our study we got 19% accuracy for lstm trained on the national corpus of Russian
% 
% [az]https://colab.research.google.com/drive/1vZ4a2T-cCpE76VGvw4E-IyODy7XapWkc Добрый день, здесь выложил часть кода, отвечающую за расчет метрики для ЛСТМ,  не могли бы сверить со своими результатами?
% 
% [ba]это довольно странное заключение: что мы считаем наиболее вероятным ответом? Что у нас является золотым стандартом вероятности ответа? Почему ответ BERT более вероятный, чем ответ людей?
% 
% [bb]why? what are the acceptability criteria?
% 
% [bc]what does the heatmap reflect?
% 
% [bd]why did you decide to do this?
% 
% [be]this is a really strange result
% 
% Can you double-check it?
% 
% [bf]and it contradicts your calculations below (3)
% 
% [bg]Непонятно, как PoS accuracy вообще может быть ниже word accuracy: если модель правильно предсказала слово, то она и его часть речи правильно предсказала.
% 
% [bh]this is also very strange
% 
% Are you absolutely sure about these results? How can you explain them?
% 
% [bi]should be aligned with the text
